openTSNE: a modular \proglang{Python} library for t-SNE dimensionality reduction and embedding

### 2023-05-23 2023-05-revision-1/main.tex
paper: add scRNA-seq pipeline to appendix

>>>>

\newpage

\begin{appendix}

\section{Single-cell RNA-seq Data Preprocessing}

We here describe the preprocessing steps applied to the single-cell RNA-seq data sets used in Fig.~\ref{fig:macosko}~\citep{macosko2015highly}, Fig.~\ref{fig:tasic}~\citep{tasic2018shared}, Fig.~\ref{fig:cao}~\citep{cao2019single}, and Fig.~\ref{fig:transform}~\citep{hochgerner2018conserved}. The standard single-cell preprocessing pipeline consists of gene filtering, gene selection, data normalization, and dimensionality reduction.

First, we discard genes detected fewer than ten times, then select the 3,000 most highly-variable genes following the gene-selection procedure from \citet{kobak2019art}. We then perform counts-per-median library-size normalization followed by log normalization and standardization without zero-centering. Finally, we extract the top 50 principal components, which are then used to construct subsequent t-SNE embeddings. This pipeline is applied to all the aforementioned data sets with the following exceptions, which serve to better illustrate our points:
\begin{itemize}
  \item In \citet{macosko2015highly}, we perform library-size normalization by computing counts-per-million and apply centering during standardization.
  \item In \citet{hochgerner2018conserved}, we select the 1,000 most variable genes.
\end{itemize}

When embedding additional data points into the existing embedding in Figs.~\ref{fig:macosko}.c and \ref{fig:transform}, we follow the procedure from \citet{policar2021embedding}. Therefore, no preprocessing is applied to the data sets from \citet{shekhar2016comprehensive} and \citet{harris2018classes}, and the similarities between the reference data set and the new data set are computed on the raw counts.

\end{appendix}
